---
title: "R Notebook"
output: html_notebook
---
install data manipulator package
```{r}
install.packages("funModeling")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("gridExtra")
install.packages("psych")
```
Loading core library
```{r}
library(tidyverse)
library(ggplot2)
library(grid)
library(gridExtra)
library(funModeling)
library(scales)
library(psych)
```

Titanic EDA


read data
data set is store under data directory
```{r}
dataset <- read.csv(file = "../data/train.csv")
```

let's explore our dataset
```{r}

head(dataset)

```

insight from r 
```{r}
dataset
metadata <-status(dataset)
write.table(metadata, file = "metadata.csv", sep = ",", row.names = FALSE) 
```
Summary of the data
```{r}
type <- c("int","int","int","chr","chr","numeric","int","int","chr","numeric","chr","chr")
description <- c("Passenger identification number","is Survived","Ticket Class","Passenger Name","Sex","Age in years","No. of sibling/spouses aboard the Tinatic","No. of parents, children aboard the Titanic","Ticket Number","Passenger fare","Cabin number","Port of Embarkation")
Key <- c("","0 = NO, 1= YES ","1 = 1st, 2 = 2nd, 3 = 3rd","","","","","","","","","C=Cherbourg ,\nQ=Queenstown,\nS=Southampton")
meta_data <- matrix(c(colnames(dataset),type,description,Key),nrow= 12, ncol=4)
colnames(meta_data) <- c("Feature Name","Data Type","Definition","Key")
grid.newpage()
grid.table(meta_data)

```



begins by tidy up the data



Removing Descriptive data with no use in this analysis(name,Ticket number, cabin, and ID)
```{r}
#Remove descriptive cols and store as dataset_exclude_Des
dataset_exclude_Des <- select(dataset, -Name, -Ticket, -Cabin, -PassengerId)
#Add Na field to "" data
dataset_exclude_Des <- dataset_exclude_Des %>% mutate_all(na_if,"")
#Remove rows with NA and store as dataset_exclude_NA
dataset_cleaned <- na.omit(dataset_exclude_Des)
#refactor the string value into group 
glimpse(dataset_cleaned)
write.table(dataset_cleaned, file = "../result/train_cleaned.csv", sep = ",", row.names = FALSE) 
```

One last time, checking that our data is good to go!
```{r}
summary(data_integrity(dataset_cleaned))

```
Now let observe the overall statisical summary of our data set

```{r}
dataset_summary <- data.frame(dataset_cleaned)
dataset_summary$Sex =  as.factor(dataset_summary $Sex)
dataset_summary$Embarked =  as.factor(dataset_summary$Embarked)
dataset_summary$Survived = as.factor(dataset_summary$Survived)
dataset_summary$Pclass = as.factor(dataset_summary$Pclass)
dataset_summary$SibSp = as.factor(dataset_summary$SibSp)
dataset_summary$Parch = as.factor(dataset_summary$Parch)
summary(dataset_summary )
```

```{r}
library(Hmisc)
dataset_quantitative <-  dataset_summary[,c("Age","Fare")]
describe(dataset_quantitative)
write.table(dataset_quantitative, file = "../result/summary_quantitative.csv", sep = ",", row.names = FALSE) 
```



```{r}
descriptive_summary <-describe(dataset_cleaned)
write.table(descriptive_summary, file = "../result/descriptive_summary.csv", sep = ",", row.names = FALSE) 
grid.newpage()
grid.table(descriptive_summary)
```





First Let's see the protortion of survival from dataset.




```{r}
##Survived
freqtable <- table(Survived = dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
##df$lab <- paste(df$Freq,paste("(",df$lab,"%)",sep=""),sep=" ")
df$lab <- paste(paste(df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Survived, Proportion, fill= Survived))
g + geom_bar(stat="identity", width = 0.5)  +scale_fill_manual(values = c("#B3000C", "#00B32C"))+  geom_text(data=df,aes(x=Survived,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Survival rate from Titanic Dataset", 

           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```
From this graph this implies  that from this dataset the survival rate  regardless to any feature is only about 40% from all passenger which is slightly less than half of all passenger aboard.


Male & Female
```{r}
freqtable <- table(Sex = dataset_cleaned$Sex)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Sex, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Sex,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Gender proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```


```{r}
freqtable <- table(SibSp = dataset_cleaned$SibSp)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( SibSp, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(SibSp,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Number of Sibling and/or spouse proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```

```{r}
freqtable <- table(Parch = dataset_cleaned$Parch)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Parch, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Parch,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Number of Parent and/or Children proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```

```{r}
freqtable <- table(Embarked = dataset_cleaned$Embarked)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Embarked, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Embarked,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Embarked proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```



Since, Survival rate is the mandatory key which we want to know, so nows on , we going to see each proportion compare to surivalrate

```{r}
freqtable <- table(Class = dataset_cleaned$Pclass)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Class, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Class,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Ticket Class proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```


  Almost half of the passenger is in 

```{r}
##Survived
freqtable <- table(Class=dataset_cleaned$Pclass,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
#get sum of each class
df$sum_class <- ave(df$Freq,df$Class,FUN=sum)

df$lab_class <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_class <- paste(paste("",df$lab_class,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( Class, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x=Class,y=Proportion,label=lab_class),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Ticket class on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```
Chance to survive of each class is significantly difference. More than half of first class passenger could survived. About half of second class passenger could survived. Lastly. Only about   
Next is Age

```{r}

hist(dataset_cleaned$Age,main= "Age Distribution of Titanic Passenger ", xlab = "Age", ylim = range(0,300))

p <- ggplot(dataset_cleaned, aes(x=Age)) + geom_histogram(fill="tomato2") +labs(title="Distribution of Passenger Age on Titanic", 
           caption="Source: Titanic Dataset(Kaggle)")
p
```
Now let see the proportion of survival from age range

```{r}
p <- ggplot(dataset_cleaned, aes(y=Age, x= factor(Survived),fill=factor(Survived)))+ scale_y_continuous("Age") +geom_boxplot()+ geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)
p
```

```{r}
p <- ggplot(dataset_cleaned, aes(x=Age,fill=factor(Survived))) + geom_histogram(position="identity", alpha= 0.5) +scale_fill_manual(values=c("#FF0000", "#00B32C")) +labs(title="Histrogram of survival& Death Distributed by Passenger Age on Titanic Data Set", 
           caption="Source: Titanic Dataset(Kaggle)")
p
```


```{r}
df <- dataset_cleaned[,c("Survived","Age")]
df_age_range <- df %>% mutate(Age_range = case_when(Age > 60   ~ '60 +',Age > 50  & Age <= 60 ~ '51 - 60',Age > 40  & Age <= 50 ~ '41 - 50',Age > 30  & Age <= 40 ~ '31 - 40',Age > 20  & Age <= 30 ~ '21 - 30',Age > 10  & Age <= 20 ~ '11 - 20',Age > 0 & Age <=10 ~ '0 - 10' ))
df_age_range

#initial
x= df_age_range$Age_range
y= df_age_range$Survived

freqtable <- table(x,Survived = df_age_range$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("    ",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge(),width=0.9,alpha=0.8)   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width=0.99), )+
      labs(title="Survival Rate in each Age range on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
```




Most of survival tend to age around 30 year
however compare all of age range, passenger with age around 0-10 have highest chance to survive


Next, Pach & Psib

```{r}
#initial axis
x= dataset_cleaned$Parch
freqtable <- table(Parch=dataset_cleaned$Parch,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
#get sum of each class
df$sum_class <- ave(df$Freq,df$Parch,FUN=sum)

df$lab_Parch <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_Parch <- paste(paste("   ",df$lab_Parch,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( Parch, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x=Parch,y=Proportion,label=lab_Parch),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Number of Parent/Child on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```



```{r}
#initial axis
x= dataset_cleaned$SibSp
freqtable <- table(SibSp=dataset_cleaned$SibSp,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$SibSp,FUN=sum)

df$lab_Parch <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_Parch <- paste(paste("  ",df$lab_Parch,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( SibSp, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x=SibSp,y=Proportion,label=lab_Parch),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Number of Siblings/spouses on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
```

Embarked

```{r}
x= dataset_cleaned$Embarked
freqtable <- table(x,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Embarked port on Titanic Data set", 

           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
```


```{r}
p <- ggplot(dataset_cleaned, aes(x=Fare)) + geom_histogram(fill="tomato2") + labs(title="Distribution of Fare on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") 
p
```


```{r}
p <- ggplot(dataset_cleaned, aes(y=Fare, x= factor(Survived),fill=factor(Survived)))+ scale_y_continuous("Fare") +geom_boxplot() +xlab("Survived")
p
```

```{r}
p <- ggplot(dataset_cleaned, aes(x=Fare,fill=factor(Survived))) + geom_histogram(position="identity", alpha= 0.5) +scale_fill_manual(values=c("#FF0000", "#00B32C")) +  labs(title="Histrogram of survival& Death Distributed by Fare on Titanic Data Set", 
           caption="Source: Titanic Dataset(Kaggle)") 
p
```

```{r}
df <- dataset_cleaned[,c("Survived","Fare")]
df_fare_range <- df %>% mutate(Fare_range = case_when(Fare > 200   ~ '200+',Fare > 100  & Fare <= 200 ~ '101 - 200',Fare >= 0  & Fare <= 100 ~ '0 - 100'))
df_fare_range

#initial
x= df_fare_range$Fare_range
y= df_fare_range$Survived

freqtable <- table(x,Survived = df_fare_range$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Survival Rate in each Fare range on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
 
```

```{r}
x= dataset_cleaned$Sex
freqtable <- table(x,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Gender on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```


```{r}
plot_num(dataset_cleaned)
```





column list
- Survived 
- Pclass




from the summary, we could see that from survived variable 
