---
title: "R Notebook"
output: html_notebook
---
install data manipulator package
```{r}
install.packages("funModeling")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("gridExtra")
install.packages("psych")
install.packages("arsenal")

```
Loading core library
```{r}
library(tidyverse)
library(ggplot2)
library(grid)
library(gridExtra)
library(funModeling)
library(scales)
##library(psych)
library(Hmisc)
library(arsenal)
```

Titanic EDA


read data
data set is store under data directory
```{r}
dataset <- read.csv(file = "../data/train.csv")
```

let's explore our dataset
```{r}

head(dataset)

```

insight from r 
```{r}
dataset
metadata <-status(dataset)
write.table(metadata, file = "../result/metadata.csv", sep = ",", row.names = FALSE) 
```
forming meta data table
```{r}
type <- c("int","int","int","chr","chr","numeric","int","int","chr","numeric","chr","chr")
description <- c("Passenger identification number","is Survived","Ticket Class","Passenger Name","Sex","Age in years","No. of sibling/spouses aboard the Tinatic","No. of parents, children aboard the Titanic","Ticket Number","Passenger fare","Cabin number","Port of Embarkation")
Key <- c("","0 = NO, 1= YES ","1 = 1st, 2 = 2nd, 3 = 3rd","","","","","","","","","C=Cherbourg ,\nQ=Queenstown,\nS=Southampton")
meta_data <- matrix(c(colnames(dataset),type,description,Key),nrow= 12, ncol=4)
colnames(meta_data) <- c("Feature Name","Data Type","Definition","Key")
grid.newpage()
grid.table(meta_data)

```



tidy up the data

```{r}
#Remove descriptive cols and store as dataset_exclude_Des
dataset_exclude_Des <- select(dataset, -Name, -Ticket, -Cabin, -PassengerId)
#Add Na field to "" data
dataset_exclude_Des <- dataset_exclude_Des %>% mutate_all(na_if,"")
#Remove rows with NA and store as dataset_exclude_NA
dataset_cleaned <- na.omit(dataset_exclude_Des)
#refactor the string value into group 
glimpse(dataset_cleaned)
write.table(dataset_cleaned, file = "../result/train_cleaned.csv", sep = ",", row.names = FALSE) 
```

check data integrity
```{r}
summary(data_integrity(dataset_cleaned))

```
re-factor and produce summary table

```{r}
dataset_summary <- data.frame(dataset_cleaned)
dataset_summary$Sex =  as.factor(dataset_summary $Sex)
dataset_summary$Embarked =  as.factor(dataset_summary$Embarked)
dataset_summary$Survived = as.factor(dataset_summary$Survived)
dataset_summary$Pclass = as.factor(dataset_summary$Pclass)
dataset_summary$SibSp = as.factor(dataset_summary$SibSp)
dataset_summary$Parch = as.factor(dataset_summary$Parch)
summary(dataset_summary )
```
reformat table & perform summary function to get Summary of Statistics

```{r}

summ_table <- tableby(~Survived + Pclass+ Sex + Age + SibSp + Parch + Fare + Embarked, data= dataset_summary )

summary(summ_table, text=FALSE)
```




additional information for quantitative factors

```{r}
descriptive_summary <-describe(dataset_cleaned[,c("Age","Fare")])
write.table(descriptive_summary, file = "../result/descriptive_summary.csv", sep = ",", row.names = FALSE) 
descriptive_summary
```





Survival rate proportion

```{r}
##Survived
freqtable <- table(Survived = dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
##df$lab <- paste(df$Freq,paste("(",df$lab,"%)",sep=""),sep=" ")
df$lab <- paste(paste(df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Survived, Proportion, fill= Survived))
g + geom_bar(stat="identity", width = 0.5)  +scale_fill_manual(values = c("#B3000C", "#00B32C"))+  geom_text(data=df,aes(x=Survived,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Survival rate from Titanic Dataset", 

           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```



Gender proportion
```{r}
freqtable <- table(Sex = dataset_cleaned$Sex)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Sex, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Sex,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Gender proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```

Sibling & Spouse proportion
```{r}
freqtable <- table(SibSp = dataset_cleaned$SibSp)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( SibSp, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(SibSp,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Number of Sibling and/or spouse proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```

Sibling & Parch proportion
```{r}
freqtable <- table(Parch = dataset_cleaned$Parch)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Parch, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Parch,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Number of Parent and/or Children proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```


Embarked proportion
```{r}
freqtable <- table(Embarked = dataset_cleaned$Embarked)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Embarked, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Embarked,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Embarked proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```



Tickey class proportion

```{r}
freqtable <- table(Class = dataset_cleaned$Pclass)
df <- as.data.frame.table(freqtable)
df$Proportion <- df$Freq/712
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
head(df)

theme_set(theme_classic())
g <- ggplot(df, aes( Class, Proportion))
g + geom_bar(stat="identity", width = 0.5, fill="tomato2")  +  geom_text(data=df,aes(Class,y=Proportion,label=lab),vjust=-.3) + 
      labs(title="Ticket Class proportion from Titanic Dataset", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```

Age histrogram
```{r}

hist(dataset_cleaned$Age,main= "Age Distribution of Titanic Passenger ", xlab = "Age", ylim = range(0,300))

p <- ggplot(dataset_cleaned, aes(x=Age)) + geom_histogram(fill="tomato2") +labs(title="Distribution of Passenger Age on Titanic", 
           caption="Source: Titanic Dataset(Kaggle)")
p
```


Histogram of fare

```{r}
p <- ggplot(dataset_cleaned, aes(x=Fare)) + geom_histogram(fill="tomato2") + labs(title="Distribution of Fare on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") 
p
```





Survival proportion of each ticket class

```{r}
##Survived
freqtable <- table(Class=dataset_cleaned$Pclass,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
#get sum of each class
df$sum_class <- ave(df$Freq,df$Class,FUN=sum)

df$lab_class <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_class <- paste(paste("",df$lab_class,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( Class, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x=Class,y=Proportion,label=lab_class),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Ticket class on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)

```



boxplot of age of both survivor & non-survivior 


```{r}
p <- ggplot(dataset_cleaned, aes(y=Age, x= factor(Survived),fill=factor(Survived)))+ scale_y_continuous("Age") +geom_boxplot()+ geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5)+ labs(title="Comparative Box plot: Age of both survivor & non-survivor on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)")
p
```





distribution of age of both survivor & non-survival
```{r}
p <- ggplot(dataset_cleaned, aes(x=Age,fill=factor(Survived))) + geom_histogram(position="identity", alpha= 0.5) +scale_fill_manual(values=c("#FF0000", "#00B32C")) +labs(title="Age Distribution of both survivor & non-survivor on Titanic Data Set", 
           caption="Source: Titanic Dataset(Kaggle)")
p
```



Surivial rate in each age range

```{r}

df <- dataset_cleaned[,c("Survived","Age")]
# devide age range into section
df_age_range <- df %>% mutate(Age_range = case_when(Age > 60   ~ '60 +',Age > 50  & Age <= 60 ~ '51 - 60',Age > 40  & Age <= 50 ~ '41 - 50',Age > 30  & Age <= 40 ~ '31 - 40',Age > 20  & Age <= 30 ~ '21 - 30',Age > 10  & Age <= 20 ~ '11 - 20',Age > 0 & Age <=10 ~ '0 - 10' ))
df_age_range

#initial
x= df_age_range$Age_range
y= df_age_range$Survived

freqtable <- table(x,Survived = df_age_range$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("    ",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge(),width=0.9,alpha=0.8)   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width=0.99), )+
      labs(title="Survival Rate in each Age range on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
```







 Proportion of survival in each parent and/or child value

```{r}
#initial axis
x= dataset_cleaned$Parch
freqtable <- table(Parch=dataset_cleaned$Parch,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
#get sum of each class
df$sum_class <- ave(df$Freq,df$Parch,FUN=sum)

df$lab_Parch <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_Parch <- paste(paste("   ",df$lab_Parch,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( Parch, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x=Parch,y=Proportion,label=lab_Parch),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Number of Parent/Child on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```





Proportion of survival in each Sibling and/or Spouse value

```{r}
#initial axis
x= dataset_cleaned$SibSp
freqtable <- table(SibSp=dataset_cleaned$SibSp,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$SibSp,FUN=sum)

df$lab_Parch <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_Parch <- paste(paste("  ",df$lab_Parch,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( SibSp, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x=SibSp,y=Proportion,label=lab_Parch),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Number of Siblings/spouses on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
```

Proportion of survival in each Embarked port

```{r}
x= dataset_cleaned$Embarked
freqtable <- table(x,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Embarked port on Titanic Data set", 

           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
```



boxplot of fare of both survivor & non-survivior 

```{r}
p <- ggplot(dataset_cleaned, aes(y=Fare, x= factor(Survived),fill=factor(Survived)))+ scale_y_continuous("Fare") +geom_boxplot() +xlab("Survived") + labs(title="Comparative Box plot: Fare of both survivor & non-survivor on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)")
p
```

Fair Distribution of both survivor & non-survivor

```{r}
p <- ggplot(dataset_cleaned, aes(x=Fare,fill=factor(Survived))) + geom_histogram(position="identity", alpha= 0.5) +scale_fill_manual(values=c("#FF0000", "#00B32C")) +  labs(title="Fair Distribution of both survivor & non-survivor on Titanic Data Set", 
           caption="Source: Titanic Dataset(Kaggle)") 
p
```


survivor proportion in each fare range

```{r}
df <- dataset_cleaned[,c("Survived","Fare")]
df_fare_range <- df %>% mutate(Fare_range = case_when(Fare > 200   ~ '200+',Fare > 100  & Fare <= 200 ~ '101 - 200',Fare >= 0  & Fare <= 100 ~ '0 - 100'))
df_fare_range

#initial
x= df_fare_range$Fare_range
y= df_fare_range$Survived

freqtable <- table(x,Survived = df_fare_range$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Survival Rate in each Fare range on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text(angle=65, vjust=0.6))+ylim(0,1)
 
```


survivor proportion in each gender

```{r}
x= dataset_cleaned$Sex
freqtable <- table(x,Survived=dataset_cleaned$Survived)
df <- as.data.frame.table(freqtable)
df$lab <- as.character(round(100 * df$Freq / sum(df$Freq),digit=2))
head(df)
#get sum of each class
df$sum_class <- ave(df$Freq,df$x,FUN=sum)

df$lab_cate <- as.character(round(100 * df$Freq /df$sum_class,digit=2))
df$Proportion <- df$Freq /df$sum_class
df$lab <- paste(paste("",df$lab,"%",sep=""),sep=" ")
df$lab_cate <- paste(paste("",df$lab_cate,"%",sep=""),sep=" ")

head(df)

x_label <- c("First Class", "Second Class", "Third Class")
type <- c("Death", "Alive")
theme_set(theme_classic())
g <- ggplot(df, aes( x, Proportion, fill=Survived))
g + geom_bar(stat="identity", position=position_dodge())   +scale_fill_manual(values = c( "#B3000C","#00B32C")) +
  geom_text(data=df,aes(x,y=Proportion,label=lab_cate),vjust=-.3,position=position_dodge(width= .95))+
      labs(title="Proportion of Survival in each Gender on Titanic Data set", 
           caption="Source: Titanic Dataset(Kaggle)") +
      theme(axis.text.x = element_text( vjust=0.6))+ylim(0,1)
```



